---
title: "IS605--Final-Exam—Hamski"
author: "James Hamski"
date: "December 19, 2015"
output: html_document
---

# Section 1 - Review
1)  
$$
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    2 & 1 & 5 & -9 \\
    6 & -1 & -2 & 4
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    -2 & 2 & 6 & -10 \\
    2 & 1 & 5 & -9 \\
    6 & -1 & -2 & 4
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    0 & 3 & 11 & -19 \\
    6 & -1 & -2 & 4
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    -6 & 6 & 18 & -30 \\
    2 & 1 & 5 & -9 \\
    6 & -1 & -2 & 4
\end{bmatrix}
\Rightarrow
$$
  
$$
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    0 & 3 & 11 & -19 \\
    0 & 5 & 16 & -26
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    0 & 3 & 11 & -19 \\
    0 & 5 & 16 & -26
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    0 & -15 & -55 & 95 \\
    0 & 15 & 48 & -78
\end{bmatrix}
\Rightarrow
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    0 & 3 & 11 & -19 \\
    0 & 0 & -7 & 17
\end{bmatrix}
$$
matrix rank = 3

$$
A =
\begin{bmatrix}
    1 & -1 & 3 & 5 \\
    2 & 1 & 5 & -9 \\
    6 & -1 & -2 & 4
\end{bmatrix}
$$

2)  
$$
t(A) = 
\begin{bmatrix}
    1 & 2 & 6 \\
    -1 & 1 & -1 \\
    3 & 5 & -2 \\
    -5 & -9 & 4
\end{bmatrix}
$$

3)  





7)  *Assuming that we are repeatedly sampling sets of numbers (each set is of size n) from an unknown probability density function. What can we say about the average value of each set?*

According to the Central Limit Theorem, we can say that these averages will be normal (Gaussian) distributed



```{r}

```

# Section 2 - Mini-coding Assignments

```{r}

```


# Section 3 - Mini-project 

```{r}

```

