---
title: "IS605-Assignment 2"
author: "J. Hamski"
date: "September 24, 2015"
output: html_document
---
##Problem Set 1

The least squares method utilizes geometry to minimize the distance between the column space of a matrix A and the vector b that represents the solutions to the linear system of equations.  
```{r}

A.1<- matrix(c(1,1,1,1,0,1,3,4), 4, 2)
b.1<- c(0,8,8,20)

left <-t(A.1)%*%A.1
right <-t(A.1)%*%b.1

x.hat <- solve(left)%*%right

solve.LS <-function(A,b){
  x.hat <- solve(t(A)%*%A)%*%(t(A)%*%b)
  return(x.hat)
}

(x.hat <- solve.LS(A.1,b.1))


#Check work using a known solution featured here: 
#https://www.youtube.com/watch?v=Z0wELiinNVQ
A.2<- matrix(c(1,1,1,1,1,2,3,4),4,2)
b.2<- c(1,3,3,5)
solve.LS(A.2, b.2)
```

The squared error is the square root of the sum of the squared components. This is the same as the Euclidian distance function "dist" in R. 
```{r}
(residuals.1 <-b.1 - (A.1 %*% x.hat)) 
error.1 <-dist(residuals.1)
error.1[3]
```


The next vector of b is a solution to the system of equations, therefore its residuals are zero. Note that due to floating point errors, the residuals must be rounded. 
```{r}
b.2 = c(1,5,13,17)
x.hat.2 <-solve.LS(A.1, b.2)

(residuals.2 <-b.2 - (A.1 %*% x.hat.2)) 
error.2 <-dist(zapsmall(residuals.2, digits = 0))
error.2[3]
```

Error e = b-p = [-1; 3.-5;3]. This is equal to the residuals calculated above. 
```{r}
(e <-b.1-b.2)
```

Show that error e is orthogal to P and each of the columns of A. The dot product of two orthogonal vectors is 0. 
```{r}
0 == e%*%b.2

0 == e%*%A.1[,1]

0 == e%*%A.1[,2]
```


##Problem Set 2
```{r}
library(readr)
#note I opted to open the data in Excel and save with commas as delineators. 
col.names = c("displacement", "horsepower", "weight", "acceleration", "mpg")
cars <- read_csv(file="auto-mpg.csv", col_names=col.names)

A.cars <- as.matrix(cars[,1:4])
b.cars <- as.matrix(cars[,5])

x.hat.cars <- solve.LS(A.cars, b.cars)
```
The linear equation that best models the dataset is:
mpg = -0.029(displacement)+0.158(horsepower)-0.0063(weight)+1.99(acceleration)

The error is:
```{r}
residuals.cars <- b.cars - (A.cars %*% x.hat.cars)
(error.cars<-(sqrt(residuals.cars[,1]%*%residuals.cars[,1])))
```

It struck me that these measures are a mix of large and small numbers. So, I tried the above methods but on a scaled dataset. 
```{r}
cars.scaled <- scale(cars)
A.cars.scaled <- as.matrix(cars.scaled[,1:4])
b.cars.scaled <- as.matrix(cars.scaled[,5])

x.hat.cars.scaled <- solve.LS(A.cars.scaled, b.cars.scaled)

residuals.cars.scaled <- b.cars.scaled - (A.cars.scaled %*% x.hat.cars.scaled)
(error.cars<-(sqrt(residuals.cars.scaled[,1]%*%residuals.cars.scaled[,1])))
```
This reduced the error by one order of magnitude. 
