---
title: "IS605—Assignment-13"
author: "J. Hamski"
date: "November 21, 2015"
output: html_document
---

*Write a program to compute the derivative of f(x) = x3 + 2x^2 at any value of x. Your function should take in a value of x and return back an approximation to the derivative of f(x) evaluated at that value. You should not use the analytical form of the derivative to compute it. Instead, you should compute this approximation using limits.*

```{r}
limit.deriv <- function(x, h){
  f.x <- (x**3)+(2*(x**2))
  f.x.h <-((x+h)**3)+(2*((x+h)**2))
  return((f.x.h - f.x)/h)
}

limit.deriv(3,0.000000001) #should equal 39
```

*Now, write a program to compute the area under the curve for the function 3x2 + 4x in the range x = [1, 3]. You should first split the range into many small intervals using some really small ∆x value (say 1e-6) and then compute the approximation to the area under the curve.*

```{r}
range.low = 1
range.high = 3
step = 1e-6

#note that you shouldn't include the area for the lower range, so intervals start at low.range plus one step
intervals <- seq(from=range.low+step, to=range.high, by=step)

curve.function <- function(x){
  height <- (3*(x**2))+4*x
  width <- step
  return(height*width)
}

areas <-lapply(intervals, FUN=curve.function)
(sum(unlist(areas))) #should equal 42
```

---
title: "KM"
author: "J. Hamski"
date: "November 22, 2015"
output: html_document
---

### Analytical Problem 1
$$\int\sin(x)\cos(x)$$ 

$$\int udv = uv - \int vdu$$
$$\int\sin(x)\cos(x) = uv - \int vdu$$


$$u = \sin(x)$$
$$du=\cos(x)$$

$$dv=\cos(x)$$
$$v=\sin(x)$$


$$= \sin(x)\sin(x) - \int\sin(x)\cos(x)$$
$$+\int\sin(x)\cos(x)$$
$$2\int\sin(x)\cos(x) = \sin^2(x)$$
$$\int\sin(x)\cos(x) = \frac{\sin^2(x)}{2}$$

### Analytical Problem 2

$$\int udv = uv - \int vdu$$
$$\int x^2e^2 = uv - \int vdu$$
$$u = x^2$$
$$du = 2x$$
$$dv = e^x$$
$$v = e^x$$

$$\int x^2e^x = x^2e^x - \int e^x2x$$
$$\int x^2e^x = x^2e^x - 2\int xe^x$$

Take $\int xe^x$ and perform integration by parts again:

$$u = x$$
$$du = 1$$
$$dv = e^x$$
$$v = e^x$$

$$\int xe^x = xe^x - \int e^x$$
$$\int xe^x = xe^x - e^x$$

Subsitute $xe^x - e^x$ back into the above equation to yield:
$$\int x^2e^x = x^2e^x - 2(xe^x - e^x)$$
$$\int x^2e^x = x^2e^x - 2xe^x + 2e^x$$
$$\int x^2e^x = (x^2-2x+2)e^x$$

### Analytical Problem 3

$$\frac{d}{dx} x\cos(x)= \frac{d}{dx}x \cdot \cos(x) + x \cdot \frac{d}{dx}\cos(x)$$
$$= 1\cdot \cos(x)+x\cdot(-\sin(x))$$
$$= \cos(x)-x\sin(x)$$

### Analytical Problem 4

$$\frac{d}{dx}(e^x)^4=(e^x)^4 \cdot \frac{d}{dx}x^4$$

$$=((e^x)^4)(4x^3)$$
