---
title: "Assign 1 - IS605"
author: "J. Hamski"
date: "August 28, 2015"
output: html_document
---

#Problem Set 1
```{r}
u = c(0.5, 0.5)
v = c(3, -4)
```

1) Dot Product
```{r}
u%*%v
```

2) Length
```{r}
(sqrt(u%*%u))

(sqrt(v%*%v))
```

3) Linear combination
```{r}
(3*u-2*v)
```

4) Angle
```{r}
vec.angle<-function(v,w){
  norm.v <-sqrt((v%*%v))
  norm.w <-sqrt((w%*%w))
  theta <- (v%*%w) / (norm.v*norm.w)
  theta <- acos(theta)
  return(theta)
}

#radians
(theta <- vec.angle(u,v))

#degrees
(deg <- theta*180/pi)
```

#Problem Set 2
Reference code used from Rosetta Code: http://rosettacode.org/wiki/Reduced_row_echelon_form#R
However, this script is difficult to read. It would be easier to understand, and potentially parallelize, if broken into more than one function. 
```{r}

rref <- function(m) {
  pivot <- 1
  norow <- nrow(m)
  nocolumn <- ncol(m)
  for(r in 1:norow) {
    if ( nocolumn <= pivot ) break;
    i <- r
    while( m[i,pivot] == 0 ) {
      i <- i + 1
      if ( norow == i ) {
        i <- r
        pivot <- pivot + 1
        if ( nocolumn == pivot ) return(m)
      }
    }
    trow <- m[i, ]
    m[i, ] <- m[r, ]
    m[r, ] <- trow
    m[r, ] <- m[r, ] / m[r, pivot]
    for(i in 1:norow) {
      if ( i != r )
        m[i, ] <- m[i, ] - m[r, ] * m[i, pivot]
    }
    pivot <- pivot + 1
  }
  return(m)
}

test <- matrix(c(1, 1, 3, 1,
              2, -1, 5, 2,
              -1, -2, 4, 6), 3, 4, byrow=TRUE)

print(test)
print(rref(test))
```

###My Work

Strategy: create a function for each of the following elemntary row operations, then wrap in a larger function that uses control structures to apply them. 

Elementary row operations:
I. Interchange two rows.
II. Multiply one row by a nonzero number.
III. Ad a multiple of one row to a different row.
```{r}

#test matrix -- input
v.input <- matrix(c(9,8,3), 3,1, byrow=TRUE)

m.input <- matrix(c(1, 2, 3,
              2, -1, 1,
              3, 0, -1), 3, 3, byrow=TRUE)

m <- cbind(m.input, v.input)

#subtract a row 
subtract.row <- function(m, rw.top, rw.bot){
  m[rw.bot,] <- m[rw.top,]-m[rw.bot,]
  return(m)
}

#interchange rows
interch.row<-function(m, rw1, rw2){
  m.copy<-m
  m[rw1,]<-m.copy[rw2,]
  m[rw2,]<-m.copy[rw1,]
  return(m)
}

#multiply row by the proper scalar
mult.row<-function(m, rw.scale, rw.elim, col.elim){
  scalar <- m[rw.elim,col.elim]
  m[rw.scale,] <- m[rw.scale,]*scalar
  return(m)
}


norow <- nrow(m)
nocolumn <- ncol(m)

```